% !TEX root = thesis.tex
\documentclass[thesis.tex]{subfiles} 
\begin{document}

\chapter{Conclusion}
\label{chap:conclusion}
Comb is a tool to extract the original dataset fed into the Mustache template
engine from the rendered template.
Over the course of this thesis we have made a case for why Comb fits into the
existing ecosystem of tools for web application development.
We began with a prototype that incorporated not only the dataset extraction but
also the coupling of said dataset to a client-side Model.
This coupling led to the realization that the applicability of Comb can be
greatly increased by solely focusing on the data extraction and leaving the
utilization of that data up to the developer who uses our tool.

Comb was split into two parts: the pre-parser tool and the parser running on
the client-side. This split freed us to implement the pre-parser in Haskell
and keep the client library free of dependencies.


We contrasted Comb with existing solutions and found that most of these require
the developer to switch to new template languages explicitly made for the
problem they are trying to solve. We identified this strategy as a framework
approach, which is a sound approach to the problem, but brings with it a reduced
applicability.
The framework approach is a general tendency we have discovered in
academic papers regarding web applications and template languages.
With Comb we have shown an alternate path to a solution.
We believe the results of this thesis warrant a reevaluation of the
current method of approach to the field of template languages for
web applications.
This belief is grounded in the lack of dependency requirements to the developer
for using Comb and amalgamated with our choice of mustache as the template
language, which because of its plentiful implementations in different languages
can almost be described as ``language agnostic''.



Our contribution, Comb, is a tool that enables developers to
build web applications with a more robust access to data embedded in rendered
templates. Comb does not restrict the developer to any kind of architectural
pattern and is built to be extended by other libraries and tools.

\end{document}
